Quantum Hoare Type Theory

### 2020-04-29 ms-paper/ms-paper.tex
Convert remaining symbols to listings literates

\usepackage{listings}
\lstloadlanguages{Haskell}
\lstdefinelanguage{QHaskell}{%
	language     = Haskell,
	morekeywords = {Qbit, U, C},
}
\lstnewenvironment{QCode}
{\lstset{language=QHaskell}%
	\csname lst@SetFirstLabel\endcsname}
{\csname lst@SaveFirstLabel\endcsname}
\lstset{
	basicstyle=\small\ttfamily,
	flexiblecolumns=false,
	basewidth={0.5em,0.45em},
	commentstyle=\color{dkgreen},
	keywordstyle=\bfseries\color{dkblue},
	stringstyle=\color{dkviolet},
	literate={+}{{$+$}}1 {/}{{$/$}}1 {*}{{$*$}}1 {=}{{$=$}}1
	{>}{{$>$}}1 {<}{{$<$}}1 {\\}{{$\lambda$}}1
	{\\\\}{{\char`\\\char`\\}}1
	{->}{{$\rightarrow$}}2 {>=}{{$\geq$}}2 {<-}{{$\leftarrow$}}2
	{<=}{{$\Leftarrow$}}2 {=>}{{$\Rightarrow$}}2
	{\ .}{{$\circ$}}2 {\ .\ }{{$\circ$}}2
	{>>}{{>>}}2 {>>=}{{>>=}}2
	{|}{{$\mid$}}1
	{/\\}{{$\wedge$}}1 {\\in}{{$\in$}}1
}

@@

\newcommand{\cket}[1]{{$\ket{#1}$}}
\newcommand\cin{{$\in$}}
\newcommand\cmt{{$\mapsto$}}
\newcommand\cmm{{$\multimap$}}
\newcommand\ccc{{$\circ$}}
\newcommand{\cpi}{$\Pi$}

@@

\subsubsection*{Testing Bell Pair}
\begin{Verbatim}[commandchars=\$\[\]]
testBell : {emp} (a, b) : (Bool, Bool) {emp /\ Id(a, b)}  -- a,b $cin {true, false}
         = do qa <= mkQbit false;
              applyU (H qa);
              qb <= mkQbit false;
              applyU (ifQ qa (X qb));
              (measQbit qa, measQbit qb)
\end{Verbatim}

@@

\begin{QCode}
testBell : {emp} (a, b) : (Bool, Bool) {emp /\ Id(a, b)}
              -- P0: emp
         = do qa <= mkQbit false;
              -- P1: P0 $ccc (qa $cmt state(false))
              applyU (H qa);
              -- P2: P1 $ccc (qa $cmt state({(0.5, false), (0.5, true)}))
              qb <= mkQbit false;
              -- P3: P2 $ccc (qa $cmt state(false))
              applyU (ifQ qa (X qb));
              -- P4: P3 $ccc ((qa, qb) $cmt state({(0.5, (false, false),
              --                              (0.5, (true, true))}))
                (measQbit qa, measQbit qb)
              -- P5: P4 $ccc ((qa $cmt -) $cmm emp) $ccc ((qb $cmt -) $cmm emp)
\end{QCode}
Finally, the \texttt{consq} rule should ensure that the assertion P5 can lead to the specified postcondition in the type of the program.

@@

\subsection{Hadamard basis states}
\begin{QCode}
qplus : {emp} r : Qbit {Id(r, $cket[+])}
      = do q <= mkQbit false;
           applyU (H q);
           return q

qminus : {emp} r : Qbit {Id(r, $cket[-])}
       = do q <= mkQbit true;
            applyU (H q);
            return q
\end{QCode}

\subsection{Creating entanglement}
\begin{QCode}
share : $cpi a : Qbit.
        {a $cin {$cket[+], $cket[-]}}   -- a $cin {b, c} is short for Id(a,b) \/ Id(a,c)
           b : Qbit
        {Id(a, b) /\ a $cin {$cket[0], $cket[1]}}
      = \a.do b <= mkQbit false;
              applyU (ifQ a (X b));
              return b
\end{QCode}

\subsection{Bell pair}
\begin{QCode}
bell : {emp} (a, b) : (Qbit, Qbit) {Id(a, b) /\ a $cin {$cket[0], $cket[1]}}
     = do qa <- qplus;
          qb <- share qa;
          return (qa, qb)
\end{QCode}

@@

\section{Teleportation}
\label{app:teleport}
% \todo{TODO: Refine assertions}
\subsection{Alice's circuit}
\begin{QCode}
alice : $cpi a : Qbit. $cpi e : Qbit
        {(Id(a, -) /\ entangled(e)}
           r : (Bool, Bool)
        {emp}
      = \a.\e.do applyU (ifQ a (X e));
                 applyU (H a);
                 (measQbit a, measQbit e)
\end{QCode}

\subsection{Bob's circuit}
\begin{QCode}
bob : $cpi m1 m2 : Bool. $cpi e : Qbit
      {entangled(e)}
         r : Qbit
      {(Id(r, -)}
      = \m1.\m2.\e.do if m1 then applyU (Z e) else ();
                      if m2 then applyU (X e) else ();
                      return e
\end{QCode}

\subsection{Teleport}
\begin{QCode}
teleport : $cpi q : Qbit. x : Pure.
      {Id(q, x)}
         r : Qbit
      {Id(r, x)}
      = \q.do (a, b) <- bell;
              (m1, m2) <- alice q a;
              tq <- bob m1 m2 b;
              return tq
\end{QCode}

>>>>

\usepackage{listings}
\lstloadlanguages{Haskell}
\lstdefinelanguage{QHaskell}{%
	language     = Haskell,
	morekeywords = {Qbit, U, C, Pure},
}
\lstnewenvironment{QCode}
{\lstset{language=QHaskell}%
	\csname lst@SetFirstLabel\endcsname}
{\csname lst@SaveFirstLabel\endcsname}
\lstset{
	basicstyle=\small\ttfamily,
	flexiblecolumns=false,
	basewidth={0.5em,0.45em},
	commentstyle=\color{dkgreen},
	keywordstyle=\bfseries\color{dkblue},
	stringstyle=\color{dkviolet},
	literate={+}{{$+$}}1 {/}{{$/$}}1 {*}{{$*$}}1 {=}{{$=$}}1
	{>}{{$>$}}1 {<}{{$<$}}1 {\\}{{$\lambda$}}1
	{\\\\}{{\char`\\\char`\\}}1
	{->}{{$\rightarrow$}}2 {>=}{{$\geq$}}2 {<-}{{$\leftarrow$}}2
	{<=}{{$\Leftarrow$}}2 {=>}{{$\Rightarrow$}}2
	{\ .}{{$\circ$}}2 {\ .\ }{{$\circ$}}2
	{>>}{{>>}}2 {>>=}{{>>=}}2
	{|}{{$\vert$}}1 {\\>}{{$\rangle$}}1
	{/\\}{{$\wedge$}}1 {\\/}{{$\vee$}}1
	{\\in}{{$\in$}}1 {\\o}{{$\circ$}}1 {\\Pi}{{$\Pi$}}1
	{|->}{{$\mapsto$}}2 {-o}{{$\multimap$}}2
}

@@

@@

\subsubsection*{Testing Bell Pair}
\begin{QCode}
testBell : {emp} (a, b) : (Bool, Bool) {emp /\ Id(a, b)}  -- a,b \in {true, false}
         = do qa <= mkQbit false;
              applyU (H qa);
              qb <= mkQbit false;
              applyU (ifQ qa (X qb));
              (measQbit qa, measQbit qb)
\end{QCode}

@@

\begin{QCode}
testBell : {emp} (a, b) : (Bool, Bool) {emp /\ Id(a, b)}
              -- P0: emp
         = do qa <= mkQbit false;
              -- P1: P0 \o (qa |-> state(false))
              applyU (H qa);
              -- P2: P1 \o (qa |-> state({(0.5, false), (0.5, true)}))
              qb <= mkQbit false;
              -- P3: P2 \o (qa |-> state(false))
              applyU (ifQ qa (X qb));
              -- P4: P3 \o ((qa, qb) |-> state({(0.5, (false, false),
              --                              (0.5, (true, true))}))
                (measQbit qa, measQbit qb)
              -- P5: P4 \o ((qa |-> -) -o emp) \o ((qb |-> -) -o emp)
\end{QCode}
Finally, the \texttt{consq} rule should ensure that the assertion P5 can lead to the specified postcondition in the type of the program.

@@

\subsection{Hadamard basis states}
\begin{QCode}
qplus : {emp} r : Qbit {Id(r, |+\>)}
      = do q <= mkQbit false;
           applyU (H q);
           return q

qminus : {emp} r : Qbit {Id(r, |-\>)}
       = do q <= mkQbit true;
            applyU (H q);
            return q
\end{QCode}

\subsection{Creating entanglement}
\begin{QCode}
share : \Pi a : Qbit.
        {a \in {|+\>, |-\>}}   -- a \in {b, c} is short for Id(a,b) \/ Id(a,c)
           b : Qbit
        {Id(a, b) /\ a \in {|0\>, |1\>}}
      = \a.do b <= mkQbit false;
              applyU (ifQ a (X b));
              return b
\end{QCode}

\subsection{Bell pair}
\begin{QCode}
bell : {emp} (a, b) : (Qbit, Qbit) {Id(a, b) /\ a \in {|0\>, |1\>}}
     = do qa <- qplus;
          qb <- share qa;
          return (qa, qb)
\end{QCode}

@@

\section{Teleportation}
\label{app:teleport}
% \todo{TODO: Refine assertions}
\subsection{Alice's circuit}
\begin{QCode}
alice : \Pi a : Qbit. \Pi e : Qbit
        {(Id(a, -) /\ entangled(e)}
           r : (Bool, Bool)
        {emp}
      = \a.\e.do applyU (ifQ a (X e));
                 applyU (H a);
                 (measQbit a, measQbit e)
\end{QCode}

\subsection{Bob's circuit}
\begin{QCode}
bob : \Pi m1 m2 : Bool. \Pi e : Qbit
      {entangled(e)}
         r : Qbit
      {(Id(r, -)}
      = \m1.\m2.\e.do if m1 then applyU (Z e) else ();
                      if m2 then applyU (X e) else ();
                      return e
\end{QCode}

\subsection{Teleport}
\begin{QCode}
teleport : \Pi q : Qbit. x : Pure.
      {Id(q, x)}
         r : Qbit
      {Id(r, x)}
      = \q.do (a, b) <- bell;
              (m1, m2) <- alice q a;
              tq <- bob m1 m2 b;
              return tq
\end{QCode}
